---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# htable

The goal of htable is to translate R data.frames into styled HTML tables.

## Installation

``` r
# install.packages("devtools")
devtools::install_github("fazetu/htable")
```

## Example

```{r}
library(htable)
ex <- data.frame(
  A = 1:10,
  B = 10:1,
  C = -4:5,
  D = letters[1:10],
  E = c(1, NA, 2, NA, 3, NA, 4, NA, 5, NA),
  Pct = runif(10),
  Big = 100000:100009,
  Money = c(100, 200, -100, -200, 1000.11, 2000.22, -1000.11, -2000.22, 0, 0.11),
  stringsAsFactors = FALSE
)
```

First create the HTable object in one of two ways:

```{r}
# ht <- HTable$new(ex)
ht <- htable(ex)
```

To get the HTML for this table:

```{r}
ht$render()
```

To write the HTML to a file:

```r
ht$writeLines("table.html")
```

In an RMarkdown document, include in your chunk settings `results="asis"` in order to get the actual HTML table to render. This chunk has `results="asis"`.

```{r, results="asis"}
ht$Rmd()
```

When not working in an RMarkdown file, use the `View()` method to see the rendered HTML table.

```r
ht$View()
```

### Format a column

3 types of numeric text formatting is supported:

* Percent
* Comma
* Dollar

```{r results="asis"}
ht$col_pct_fmt(6)
ht$col_comma_fmt(7)
ht$col_dollar_fmt(8)
```

Since `ht` is a reference class object each formatting operation is done in place.

```{r, results="asis"}
ht$Rmd()
```

These operations change the `contents` field of `ht`. This matrix is what is used to get the content for each HTML table cell.

### Style columns

This demonstrates many of the available column styles. All available column stylings start with the prefix "col_".

```{r, results="asis"}
ht$col_bg_color(1, "red", include_header = TRUE)
ht$col_color(1, "white", include_header = TRUE)
ht$col_bold(2)
ht$col_italic(3)
ht$col_color_scale(6)
ht$col_data_bar(8)
ht$Rmd()
```

These operations change the `styles` field of `ht`. This matrix is what is used to apply styles to each `<th>` and `<td>` tag of the HTML table.

### Reset styles and contents

To start with a clean slate of styles:

```{r, results="asis"}
ht$clear_styles()
ht$Rmd()
```

This clears the `styles` field of `ht`. There is also a similar `clear_all_styles()` method. This also clears `table_style`, `thead_style`, `tbody_style`, and `tr_styles` fields of `ht`.

To start with a clean slate of contents:

```{r, results="asis"}
ht$reset_contents()
ht$Rmd()
```

This resets the `contents` field of `ht`.

### Style rows

This demonstrates many of the avialable row and header styles. All available row styles start with the prefix "row_" and all available header styles start with "header_".

Add an alternating text and background color:

```{r, results="asis"}
# first colors apply to odd rows and second color applies to even rows
ht$row_alt_color("white", "black", include_header = TRUE)
ht$row_alt_bg_color("black", "orange", include_header = TRUE)
ht$Rmd()
```

Other row and header styles:

```{r, results="asis"}
ht$clear_styles()
ht$row_bg_color(1, "red")
ht$row_color(2, "blue")
ht$row_bold(3)
ht$row_italic(4)

ht$header_bg_color("black")
ht$header_italic()
ht$header_color("white")
ht$Rmd()
```

### Style cell

Can target an individual cell as well. All available cell styles start with the prefix "cell_".

```{r, results="asis"}
ht$clear_styles()
ht$cell_add_style(1, 1, "background-color:blue;font-weight:bold;color:white;")
ht$Rmd()
```

### Final table

This demonstrates a typical construction of a final table.

```{r, results="asis"}
ht$clear_styles()

# pleasant overall styles
ht$add_styles("padding:10px;font-size:14px;font-family:arial")
ht$table_style_add("border-collapse:collapse;")
ht$tr_styles_all_add("border-bottom:solid #A9A9A9;border-width:1px 0;")
ht$tr_styles_which_add(1, "border-width:3px 0;")
ht$tr_styles_which_add(length(ht$tr_styles), "border:none;")

# edit the contents
ht$col_pct_fmt(6)
ht$col_comma_fmt(7)
ht$col_dollar_fmt(8)

ht$col_data_bar(1, "steelblue") # specify a color for bars
ht$col_color_scale(6:8) # use default colors
ht$Rmd()
```
